---
title: "Project1"
author: "Destiny McKenzie and Lily Nathan"
format: pdf
editor: visual
---

## 
```{r}
variable_table <- data.frame(
  Variable = c(
    "US_retailPrice", "pieces", "minifigs", "weight",
    "height", "width", "depth", "agerange_min",
    "rating", "reviewCount", "theme", "themeGroup", "category"
  ),
  
  Role = c(
    "Response",
    rep("Quantitative Predictor", 9),
    rep("Categorical Predictor", 3)
  ),
  
  Data_Type = c(
    "Numeric", "Numeric", "Numeric", "Numeric", "Numeric",
    "Numeric", "Numeric", "Numeric", "Numeric", "Numeric",
    "Categorical", "Categorical", "Categorical"
  ),
  
  Description = c(
    "Original U.S. retail price of the LEGO set in U.S. dollars",
    "Total number of LEGO pieces",
    "Number of minifigures included in the set",
    "Weight of the LEGO set",
    "Height of the LEGO set",
    "Width of the LEGO set",
    "Depth of the LEGO set",
    "Minimum recommended age for the set",
    "Average user rating for the set",
    "Number of user reviews",
    "Specific LEGO theme associated with the set",
    "Broader grouping of LEGO themes",
    "Category or type of LEGO set"
  )
)
variable_table

```
```{r}
variables <- c(
  "US_retailPrice",
  "pieces",
  "minifigs",
  "weight",
  "height",
  "width",
  "depth",
  "agerange_min",
  "rating",
  "reviewCount",
  "theme",
  "themeGroup",
  "category"
)

missing_table <- data.frame(
  Variable = variables,
  Missing = colSums(is.na(data[variables]))
)

missing_table
```
```{r}
variables <- c(
  "US_retailPrice",
  "pieces",
  "minifigs",
  "weight",
  "height",
  "width",
  "depth",
  "agerange_min",
  "rating",
  "reviewCount",
  "theme",
  "themeGroup",
  "category"
)

data_complete <- data[complete.cases(data[variables]), ]
nrow(data)
nrow(data_complete)
colSums(is.na(data_complete[variables]))
```
Once identifying the predictor variables for our research question, we created a table that would look into how many missing values each predictor variable had. After identifying how many data points had missing values, we created a new data set, named data_complete, containing only lego sets with complete predictor variable values. This allows us to make consistent comparisons between sets and prevents missing values from affecting our statistical analyses or limiting the observations available for certain calculations. 
